### 2024-06-26 thesis/ms-fletcher-porter.tex
Add new OSI model figure

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/osi_model}
  \caption{The OSI model understands network communication to be a layer cake of abstractions over the physical communication media.  Taken from~\cite[§6.1.3]{ISO7498-1}.}\label{fig:osi_model}
\end{figure}

@@

When these tokens are combined into phrases, that's syntax. Syntax is the structure into which meaning is applied. The \verb|printf| example is syntactically parsed by C compilers as in figure~\ref{fig:printf_syntax}. The syntax tells that the postfix expression represented by the identifier printf has the argument list of \verb|"I have %u cows\n"| (a string literal) and \verb|num_cows| (an identifier). What any of these tokens mean is the domain of semantics.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{assets/printf_syntax}
  \caption{The above printf example with syntax labeled according to K{\&}R C~\cite{Kernighan1978}}\label{fig:printf_syntax}
\end{figure}

@@

An FPGA, like a CPU, is a programmable chip that performs computations.  However, they differ in when the program is read.  In a CPU, a program is read and executed as the program runs.  A CPU reads an instruction from an address in memory specified by an instruction counter internal to the CPU.  That instruction will be loaded into the CPU and in executing it, the CPU will change its state, perhaps by loading a value from memory into a CPU register, executing an arithmetic or logic instruction, interfacing with IO, or quite often incrementing the instruction counter.  An FPGA, on the other hand, reads and loads the entire program before execution begins.  It does this by mutating its internal circuits to implement the program.  With a lower-level program implementation, \ie a hardware implementation rather than software, an FPGA is able to see higher performance for similar programs.  Instead of indirection through memory lookups, instruction counters, bus loads, and registers, an FPGA can pipeline all necessary instructions in a row.  A PI controller can be implemented as a subtraction circuit, followed by two gains and an accumuluator circuit, followed by an addition circuit.  The circuit would be essentially passive in this case, though there can be memory for more complex logic like would be needed for PD controllers.

@@

It also tested this theory against several case studies.  In one case, a hydraulic steering linkage testbed showed that a change in one absraction layer can demand rearchitecture of the other abstraction layers if the control system is not carefully designed.  In another, an organization employed variations on one tech stack to build robots based on soft actuators, presumably to capitalize on institutional knowledge of that stack.  The last used the development of a stewart platform to show how a control system can be developed based on this theory.  By organizing the tools used in these case studies based on the function they perform and the interface they provide, I generated these compelling insights into the architecture of the control system.

@@

This theory can also be extended.  Throughout this work, it has been called a \textit{mechatronic} tech stack, but truly it's a control system tech stack.  To make it a true mechatronic tech stack, it should include tooling for mechanical and electronic systems.  A mechanical section of the stack would likely grow from the bottom of the stack shown in figure~\ref{fig:mechatronic_tech_stack}, not colliding with the current stack, but an electrical component would likely be more complicated, since it is needed by both the mechanical stack for transducers and sensors, and by the control stack since the computer \textit{is} an electrical component, and the IO layer often needs some elecrical finesse, as noted in §\ref{sec:stewart_platform}.  It could also be extended in the software realm; the production of graphical monitoring systems and tele-operation programs generally require substantial non-control software which could be described by different layers, perhaps replacing the control layer in the present stack.

This theory would also benefit from the study of adjacent work that doesn't develop the mechatronic tech stack itself.  This theory describes the function of mechatronic tooling, but it provides no notion of what tools are best for a particular application.  In evaluating the case studies, some initial guesses were developed, like that tools should encourage loose coupling, take advantage of institutional knowledge, and be easy to troubleshoot.  This can be formalized by interviewing practitioners and querying what qualities they value in tools as well as by studying extant literature on the subject of tool quality.

>>>>

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/osi_model}
  \caption{The OSI model understands network communication to be a layer cake of abstractions over the physical communication media.  An individual tool may implement the behviours and interfaces of of one or multiple layers.  The layering enables design freedom in how a network is implemented.  Communication over HTTP usually goes over a TCP transport through an IP network and to other destinations over WiFi or Ethernet, but Lindgren et al.\ show that it's possible to send IP signals over CAN~\cite{Lindgren2008}, and Ethernet over EtherCAT is possible as well~\cite[§1.9.3]{EtherCATFieldbus}.  High level protocols can also be replaced to obtain a lower level of abstraction for performance or design flexibility, as was done for the video game \emph{DOOM} in 1993, as well as for contemporary game engines like Godot or Unity~\cite{DOOM1993,GodotMultiplayer,UnityTransport}.}\label{fig:osi_model}
\end{figure}

@@

When these tokens are combined into phrases, that's syntax. Syntax is the structure into which meaning is applied. The \verb|printf| example is syntactically parsed by C compilers as in figure~\ref{fig:printf_syntax}. The syntax tells that the postfix expression represented by the identifier \verb|printf| has the argument list of \verb|"I have %u cows\n"| (a string literal) and \verb|num_cows| (an identifier). What any of these tokens mean is the domain of semantics.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/printf_syntax}
  \caption{The above printf example with syntax labeled according to K{\&}R C~\cite{Kernighan1978}}\label{fig:printf_syntax}
\end{figure}

@@

An FPGA, like a CPU, is a programmable chip that performs computations.  However, they differ in when the program is read.  In a CPU, a program is read and executed as the program runs.  It reads an instruction from an address in memory specified by an instruction counter internal to the CPU.  That instruction will be loaded into the CPU and in executing it, the CPU will change its state, perhaps by loading a value from memory into a CPU register, executing an arithmetic or logic instruction, interfacing with IO, or any of myriad other things.  An FPGA, on the other hand, reads and loads the entire program before execution begins.  It does this by mutating its internal hardware circuits to implement the program.  With this lower-level program implementation an FPGA is able to see higher performance for similar programs.  Instead of indirection through memory lookups, instruction counters, bus loads, and registers, an FPGA can pipeline all necessary instructions in a row.  A PI controller can be implemented as a subtraction circuit, followed by two gains and an accumuluator circuit, followed by an addition circuit.  The circuit would be essentially passive in this case, though there can be memory for more complex logic like would be needed for a PD controller.  The programs for FPGAs are generally written in hardware description languages (HDLs) like Verilog or VHDL which have important hardware concepts like time core to the language, though there are libraries for conventional languages like C++ or Rust to generate HDL code.

The Solid Mechanics Lab employed both a CPU and an FPGA for their control.  The FPGA ran the ``hot'', or high-frequency, calculations, namely the PID loop and also did analog input from the analog-digital converter (ADC).  The ADC actually sampled signals at $\qty{100}{\kilo\hertz}$ and the FPGA program was made to super-sample the input to get an $\qty{18.3}{\bit}$ signal out of a $\qty{16}{\bit}$ ADC.

@@

It also tested this theory against several case studies.  In one case, a hydraulic steering linkage testbed showed that a change in one absraction layer can demand rearchitecture of the other abstraction layers if the control system is not carefully designed.  In another, an organization employed variations on one tech stack to build robots based on soft actuators, presumably to capitalize on institutional knowledge of that stack.  Another used the development of a stewart platform to show how a control system can be developed based on this theory.  The last showed how performance requirements for a strain testing machine can drive the architecture of a control system and how that can be captured by this model.  By organizing the tools used in these case studies based on the function they perform and the interface they provide, I generated these compelling insights into the architecture of the control system.

@@

This theory can also be extended.  Throughout this work, it has been called a \textit{mechatronic} tech stack, but truly it's a control system tech stack.  To make it a true mechatronic tech stack, it should include tooling for mechanical and electronic systems.  A mechanical section of the stack would likely grow from the bottom of the stack shown in figure~\ref{fig:mechatronic_tech_stack}, not colliding with the current stack, but an electrical component would likely be more complicated, since it is needed by both the mechanical stack for transducers and sensors, and by the control stack since the computer \textit{is} an electrical component, and the IO layer often needs some elecrical finesse, as noted in §\ref{sec:stewart_platform}.  It could also be extended in the software realm; the production of graphical monitoring systems and tele-operation programs generally require substantial non-control software which should be described by different layers, perhaps branching away from the control layer to have a parallel monitoring layer.

This theory would also benefit from the study of adjacent work that doesn't develop the mechatronic tech stack itself.  This theory describes the function of mechatronic tooling, but it provides no notion of what tools are best for a particular application.  In evaluating the case studies, some initial guesses were developed, like that tools should encourage loose coupling, take advantage of institutional knowledge, be easy to troubleshoot, and meet certain performance requirments.  This can be formalized by interviewing practitioners and querying what qualities they value in tools as well as by studying extant literature on the subject of tool quality.  These studies would be synthesized into a coherant model for representing the quality of a tool.  Likely this would manifest as a multi-dimensional which a practitioner would interpret to get the best result as apposed to a one-dimensional quality rating that gives practitioners no agency.
